---
title: "example_simulation"
output: html_document
---

This document contains a full example of how to source and run the models to generate data. Refer to the [README](./01_model_scripts/README.md) in the `01_model_scripts` folder for additional information on using the models, the model purpose, and model assumptions. Be sure to update the README in the `02_run_models` folder, found [here](./02_run_models/README.md), each time you run the model to generate new data.  


## Step 1 - Packages

Load the packages required to run the model  

```{r packages, message=F, warning=F}
library(tidyverse)
library(purrr)
library(here)
```


## Step 2 - Source Model

Source the model from the `01_model_scripts` folder, we will be doing this using the *here* package which sets the file path to the working directory to wherever you have it stored on your local computer. You can find out your local file path by typing here() in the console       

```{r source-model}
# In this example we will be using the sim_fishery() model so this is the model we want to source
source(file = here::here("01_model_scripts/sim_fishery.R"))
```

## Step 3 - Inputs

The model requires 9 inputs which are defined in the R script and a specified assessment year interval.  

**Assessment Interval**  
This is the interval with which the model will be making decisions and we need to set it up as a numeric vector. A few important notes about the assessment interval:  

 - The model requires it to be named "assess_int"  
 - The model will always make a decision in year 1 and you don't need to specify this in your interval  
 - The first number and the last number should match; the first number is the first year a decision is made after Year 1, and the last number is the interval  
 - The middle number is the length of the simulation, this should match what you use as your *years* parameter input, the default is 100 years  
 
In our example we will use a 10 year assessment interval  

```{r define-assessment-interval}

# Make decision starting in year 10 every 10 years for a total of 100 years
assess_int <- seq(10,100,10) 
```

**Model Inputs**
The model inputs are described in the R script. While you can simply use the notation sim_fishery(b=4000, r=0.6, r_s = -0.00767 ... hcr = 0.85) this will only run the model a single time for a single set of specifications. This may be useful in some situations and is one method for generating data.  

We have chosen to run the model simultaneously over a list of inputs to represent a range of different input values. This allowed us to compare the outcomes across some of the different paces of climate change, growth rates, initial biomass conditions, choices for harvest control rules etc. Examples for the `r_s` input can be found in the [climate_scenarios](./reference/climate_scenarios.md) file in the `references` folder. This is a sample code for how to create list of simulations, the values for each input can be altered depending on what is of interest to you. 

Our list includes a range of values for the following input parameters:  

 - `b`: 1500, 4000, 6000  represent overfished (.15K), fully fished (0.4K), and healthy stocks (0.6K)  
 - `r`: use a sequence of growth rates from 0.1 - 0.8 to reflect slower growing and faster growing species  
 - `r_s`: sequence of climate change paces from - 0.01767 to + 0.01623
 for examples refer to the [climate_scenarios](./reference/climate_scenarios.md) file in the `references` folder  
 - `hcr`: three discrete harvest control rules to compare 10%, 20%, and 30% reductions  
 

```{r experiment-inputs}

# Create a list with ranges you want to test for each of the required model inputs

list_inputs <- list(
  
  b     = c(1500, 4000, 6000), 
  r     = seq(0.1, .8, .1), 
  r_s   = seq(-0.01767, 0.01623, 0.002),  
  r_p_s = -0.00567,
  error = 0.1, 
  p     = 0.2, 
  k     = 10000, 
  years = 100,
  hcr   = c(0.7, 0.8, 0.9) 
)

# Generate all combinations to allow for a full comparison across groups
# Note if this list size gets really large the model will run very slowly
inputs <- list_inputs %>%
  cross()


```

## Step 4 - Running the Model

As we mentioned there are two ways to run the model. The first is simply using a single value for each input, this method is shown in the first code chunk. The second is using the list of inputs that we created and running the model over each individual list, this method is shown in the second code chunk.  

**Method 1**

This method uses a single value for each model parameter  
Note: if you don't want to change the values of p, k, or years from the default you don't need to include them  
Default values are given in the R script  

```{r run-method1}

run1 <- sim_fishery(b     = 4000,
                    r     = 0.6,
                    r_s   = -0.00767,
                    r_p_s = -0.00567,
                    error = 0.1,
                    hcr   = 0.85)
```

Output is a dataframe with 11 columns and 100 rows (years of the simulation), an explanation of each column variable is given in the R script and the README file in this folder.  


**Method 2**

This method uses *lapply()* to apply the model over the list of inputs generated. Again, we don't have to include the p, k, or years inputs if we are using the default values.  

```{r run-method2}

run2 <- lapply(1:length(inputs), function(x) 
  sim_fishery(b     = inputs[[x]]$b,
              r     = inputs[[x]]$r, 
              r_s   = inputs[[x]]$r_s,
              r_p_s = inputs[[x]]$r_p_s,
              error = inputs[[x]]$error,
              hcr   = inputs[[x]]$hcr))
```

Output here is a list of 1224 items, one for each simulation, with each list containing the same 11 column by 100 row data frame with the results from that specific model run   

## Step 5 - Add Identifiers  

We have chosen to add some identfiers to our data to make data analysis easier. This step is optional although we *highly recommend* it. When the results are unlisted into a single csv file we lose some of the ability to see what inputs went into the model run and these identifiers can be helpful for differentiating between simulations. We recommend including at a minimum the ID, r_s, r_p_s, and hcr columns. If other inputs, for example error, were varied in your simulations you may want to add those as well. These identifer columns can be used to compare across groups and are helpful for graphing results.  

```{r add-identifiers}
# Create the list of column names (cols 1-11 are the model results)
# Model outputs are described in the R script and the README in this folder
# Include names for the columns we are going to add as identifiers (in this case id - b_0) 

col_names <- c("b", "c", "year", "r",
               "r_p", "f", "f_msy",
               "f_msy_p", "f_ratio",
               "f_ratio_p", "f_ratio_err",
               # Added identifier columns
               "id", "r_0","r_s", "r_p_s",
               "hcr", "b_0") 

# Store the updated simulations with added identifier columns in a new list
results_id <- list()

# Add the identifiers to each experiment in the list
for (i in 1:length(run2)) {
  
  # Unlisting the results to add identifier columns
  dat <- as.data.frame(matrix(unlist(run2[i]), ncol = 11)) 
  
  # Add simulation ID to keep track of different runs
  dat$id <- i 
  
  # Add identifiers from the original input list
  dat$r_0   <- rep(inputs[[i]]$r, 100) 
  dat$r_s   <- rep(inputs[[i]]$r_s, 100)
  dat$r_p_s <- rep(inputs[[i]]$r_p_s, 100) 
  dat$hcr   <- rep(inputs[[i]]$hcr, 100) 
  dat$b_0   <- rep(inputs[[i]]$b, 100) 
  
  
  results_id[[i]] <- dat 
}

# We want to have a single tidy format file (not a list)
results_all <- bind_rows(results_id)
colnames(results_all) <- col_names
```

## Step 6 - Save Results

The results should be saved to prevent the need to re-run the model every time as well as to preserve reproducible results. The model involves some randomness so the exact results will never be the same from one run to the next. Therefore, saving your results preserves the outcomes from any given model run. The results should be saved to the `03_model_output` folder and we save them as csv files.   

Again we use the *here* package to assist with file paths and make sure results are saved to the correct location.  

Recommendations for naming conventions can be found in the [SOP](./reference/SOP_data_management.md) style guide, for this example we will call this "example_results.csv"

```{r save-csv}

write.csv(results_all, file = file.path(here::here(),"/03_model_output/example_results.csv"),
           row.names = FALSE)
```

