---
title: "example_analysis"
output: html_document
---

This document provides an example of generating a graph to compare outcomes of fishery health in year 100 based on different harvest control rules.    

## Step 1 - Packages

We are using the *tidyverse* to wrangle our data for graphing and we will use the *here* package for setting the files path to the data. If you have not installed these packages, you must first do that using install.packages("package-name")
```{r packages, message=F, warning=F}
library(tidyverse)
library(here)
```


## Step 2 - Read Data

Read in data from the `05_analysis_data` folder 

```{r read-data, message=FALSE, warning=FALSE}

ex_df <- read_csv(file=file.path(here::here(),"/05_analysis_data/ex_analysis_data.csv")) 

```


## Step 3 - Wrangle for Graphing

The graphs in our project's report have a proportion of healthy fisheries on the y-axis and the categories for the action of interest on the x-axis. For this graph, our action of interest are the different harvest control rules (hcrs) we used when we ran the model. We can calculate a proportion of healthy fisheries using the "status" identifier we added during the processing code.  

The data wrangling to create this graph has three parts: 

 * Filter for only negative climate change (depending on your interest)
 * Group by your action of interest (in this case the hcr)  
 * Count the number of simulations in the different status categories (in this case we just have "over" and "healthy")  
 * Use the spread() function so we can get the count for over and healthy as columns have only three rows (one for each hcr)  
 * Find the total number of simulations 
 * Find the proportion of simulations with a status of "healthy"  
 
```{r wrangle-graphing-data}
# Create a data frame we can use to graph proportion of healthy fisheries for each of our hcrs
hcr_props <- ex_df %>% 
  filter(r_s < 0) %>% 
  group_by(hcr) %>% 
  count(status) %>% 
  # Want to have a 3 row by 3 column df for making this graph 
  spread(key = "status", value = "n") %>% 
  mutate(simulations = sum(over + good)) %>% 
  mutate(prop_good = good / simulations)
```


## Step 4 - Graph

Create a bar graph with the hcr's as the x-axis and the proportion of healthy fisheries as the y-axis. This allows us to compare the impact of using a different harvest control rule on the over proportion of healthy fisheries.  

```{r graph}
hcr_graph <- ggplot(hcr_props, aes(x = hcr, y = prop_good)) +
  geom_col(alpha = 0.8, fill = "#079EDF") + 
  theme_light()+
  coord_cartesian(ylim=c(0,1), expand = FALSE ) +
  # Show lower fishing reduction on the left and higher fishing reduction on the right
  # Use the % reduction instead of the decimal as the axis label
  scale_x_reverse(breaks=c(0.7,0.8,0.9), labels=c("0.7" = "30%","0.8" = "20%","0.9" = "10%")) +
  labs(title = "Fishing Reduction", x = "Reduction Amount (%)", y = "Proportion Healthy") + 
  # General graph syling for all figures in the report
  # Styling includes: x/y axis tick marks, x/y axis labels
  # plot title, legend, grid lines, and background transparency
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=15),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 20), 
        axis.title.x = element_text(face = "bold", size = 15), 
        axis.title.y = element_text(face = "bold", size = 15), 
        legend.title.align=0.5, panel.background = element_rect(fill = "transparent",colour = NA), 
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.title=element_text(size=15), legend.text=element_text(size=12), 
        legend.key = element_rect(fill = "transparent", colour = "transparent"), 
        legend.background = element_blank(), legend.box.background = element_blank(),
        panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) 

hcr_graph
```


## Step 5 - Save as PNG

Save the graph as a .png file to the `07_figures` folder. We will do this using the ggsave() function. Recommendations for naming conventions can be found in the [SOP](./Reference/SOP_data_management.md) style guide, for this example we will save this figure as "example_figure.png".    

Specifying the height and width of the figure and a transparent background are helpful when making figures for reports or presentations.  

```{r save-graph}
ggsave(hcr_graph,
       filename = file.path(here::here(),"07_figures/example_figure.png"), 
       bg = "transparent", 
       width = 6, height = 4) 
```

